01 - TreehopperFitness: Quality control

### 2023-08-11 scripts/Publilia-01-qc.Rmd
Moved all data prep for `Publilia-03-survivorship.Rmd` to `Publilia-01-qc.Rmd`

# 6 - Write data to `.rds` for further analysis in clean environment
```{r}
write_rds(All_data_v6, here("saved_rds/All_data_v6.rds"))
write_rds(clutch_status_guarding_summary_v2, here("saved_rds/clutch_status_guarding_summary_v2.rds"))
write_rds(last_day_fID, here("saved_rds/last_day_fID.rds"))
```

>>>>

# 6 - Write data for `Publilia-02-strats.Rmd` to `.rds` for further analysis in clean environment

## `Publilia-02-strats.Rmd`
```{r}
write_rds(All_data_v6, here("saved_rds/All_data_v6.rds"))
write_rds(clutch_status_guarding_summary_v2, here("saved_rds/clutch_status_guarding_summary_v2.rds"))
write_rds(last_day_fID, here("saved_rds/last_day_fID.rds"))
```

***

*Prep data for `Publilia-03-survivorship.Rmd`:*

# 7 - Assigning nymph origins

Nymphs migrate away from their natal leaf. On plants with multiple clutches, it was sometimes possible to determine the origin of nymphs later in the season because sometimes groups of nymph don't travel very far from their natal leaf and mix with other groups. I manually curated examples of this, so that clutch-level stats on hatching success and nymph survivorship could be determined on the plants with multiple clutches. Some clutches (as a rule only on multi-clutch plants) did not have a clear record of nymphs that emigrated from the natal leaf and thus could not be included in future clutch-level analyses of survivorship.

## 7.1 - Adding second gen adult counts per leaf to clutch status stats
```{r}
# counting adults
eA_status_v1 <- All_data_v6 %>%
  group_by(row_number()) %>% # sum of newly exclosed adults
  filter(any(!is.na(c(New_eA, Old_eA)))) %>%
  mutate(n_newA = ifelse(is.na(New_eA),
                         NA,
                         sum(as.numeric(regmatches(New_eA, gregexpr("[0-9]+", New_eA))[[1]]))),
         n_oldA = ifelse(is.na(Old_eA),
                         NA,
                         sum(as.numeric(regmatches(Old_eA, gregexpr("[0-9]+", Old_eA))[[1]]))),
         n_A = sum(n_newA, n_oldA, na.rm = TRUE)) %>%
  ungroup() %>%
  select(Date, location, n_newA, n_oldA, n_A)

All_data_v6_clutch_status_stats_v2 <- All_data_v6 %>%
  # add egg measures, clutch status stats (including initiation dates, hatch dates, egg fate, etc)
  left_join(clutch_status_v2, by = c("location", "Date")) %>%
  left_join(clutch_status_stats_v1, by = c("location")) %>%
  left_join(eA_status_v1, by = c("Date", "location")) %>%
  left_join(select(clutch_status_guarding_summary_v2, location, never_hatched)) %>%
  mutate(N1 = as.numeric(N1),
         N2 = as.numeric(N2),
         N5 = as.numeric(N5))
```

## 7.2 - Manually check nymph origins for revision
First create a table of all entried from clutches that were not flagged in the field as having ambiguous origin (e.g. "leaf 6 or 9") in cases were more than one clutch was on the plant. These were revised manually to specify a leaf of origin for the nymphs, where possible.
```{r}
NOr_status_resolveable_all_unambig_clutches <- All_data_v6_clutch_status_stats_v2 %>% 
  select(Date, Dorm, Plant, Leaf, location, 
         egg_num, N1, N2, N5, n_A, nymph_origin, initiation_date, hatch_date,
         any_NOr_ambig_loc, Omit_for_abandon_time, never_hatched,
         n_clutches_on_plant) %>%
  # keep only entries with some nymphs or adults on a leaf
  # OR entries for leaves where clutches have been initiated
  group_by(row_number()) %>%
  filter(!is.na(Leaf) & any(N1 > 0, N2 > 0, N5 > 0, n_A > 0) | 
           Date >= initiation_date) %>% # nrow = 17944
  # flag entries where the nymph origin is text (typically ambiguous, e.g. "8 or 9")
  mutate(NOr_ambig_entry = !is.na(nymph_origin) & is.na(as.numeric(nymph_origin))) %>% 
  ungroup() %>%
  # get rid of duplicate rows from multiple females in the same location
  distinct(Date, location, .keep_all = TRUE) %>% # nrow = 15674
  group_by(Dorm, Plant) %>%
    # filter out any plants with a clutch that was omitted for abandon time
    filter(!any(!is.na(Omit_for_abandon_time))) %>%
    # flag plants with any ambiguous entries (e.g. "8 or 9") skip for checking
    mutate(NOr_ambig_plant = any(NOr_ambig_entry)) %>%
  ungroup()

NOr_status_resolveable_mult_clutches <- NOr_status_resolveable_all_unambig_clutches %>% # nrow = 15674
  group_by(Dorm, Plant) %>%
    # plants with mutliple clutches only
    filter(any(n_clutches_on_plant > 1)) %>%
  ungroup() %>%
  group_by(row_number()) %>%
    # flag for manual checking entries with any nymphs/adults, missing nymph origin, 
    # but not flagged for ambiguous origin in the field.
    # these are the most likely to be easily resolved with minimal work
    mutate(revise_NOr = ifelse(any(N1 > 0, N2 > 0, N5 > 0, n_A > 0) &
             !NOr_ambig_plant,
           "MANUAL CHECK",
           NA)) %>% 
  ungroup()

NOr_manual_check_table <- NOr_status_resolveable_mult_clutches %>%
  group_by(Dorm, Plant) %>%
    filter(any(revise_NOr == "MANUAL CHECK")) %>% # nrow = 2830
  ungroup() %>%
    group_by(row_number()) %>%
    filter(any(N1 > 0, N2 > 0, N5 > 0, n_A > 0)) %>%
  ungroup() %>%
  select(Date, Dorm, Plant, Leaf, location, egg_num, N1, N2, N5, n_A, nymph_origin,
         initiation_date, hatch_date,
         revise_NOr) %>%
  arrange(Dorm, as.numeric(Plant), as.numeric(Leaf), Date)
# 83 clutches on 34 plants to check

# write_csv(NOr_manual_check_table %>% filter(revise_NOr == "MANUAL CHECK"), "NOr_manual_check_tmp.csv")

NOr_manual_revisions <- read_xlsx(in_filename, sheet = "NOr_manual_check", col_types = c("date", "text", "text", "text", "text", "text", "text", "text", "text", "text", "text", "text", "text", "text", "logical", "text")) %>%
  mutate(Date = as_date(Date),
         NOr_rev_loc_manual = paste0(Dorm, Plant, "_", NOr_revised)) %>%
  filter(!cant_resolve) %>%
  select(Date, location, NOr_rev_loc_manual, NOr_revised_manual = NOr_revised)

NOr_manual_revisions %>%
  distinct(NOr_rev_loc_manual) # 43/83 initiated clutches (43/59 hatched) with successfully revised NOr
```

## 7.3 - Implement revisions by creating the new variables NOr_revised and NOr_rev_loc
```{r}
# if nymph origin was clearly resolveable, the revisions were made in NOr_manual_revisions
# updating NOr_status with revisions so that it can be left_joined to other data and
# counts of nymphs and adults on NOr can be calculated
NOr_status_revised_all_unambig_clutches <- NOr_status_resolveable_all_unambig_clutches %>%
  left_join(NOr_manual_revisions, by = c("Date", "location")) %>%
  group_by(Dorm, Plant) %>%
    mutate(lone_hatcher_leaf = ifelse(length(unique(Leaf[!is.na(initiation_date)])) == 1,
                                 unique(Leaf[!is.na(initiation_date)]),
                                 NA),
           NOr_revised = ifelse(!is.na(NOr_revised_manual),
                                       NOr_revised_manual,
                                ifelse(any(!is.na(lone_hatcher_leaf)),
                                       lone_hatcher_leaf,
                                       # if no clutches hatched on the plant, NOr_revised = Leaf
                                       # for leaves with eggs
                                       ifelse(!any(!is.na(hatch_date)) & !is.na(egg_num),
                                              Leaf,
                                              NA)))) %>%
  ungroup() %>%  
  group_by(location) %>%
    mutate(NOr_revised = ifelse(any(NOr_revised == Leaf) & is.na(NOr_revised),
                                Leaf,#unique(NOr_revised[!is.na(NOr_revised)]),
                                NOr_revised)) %>%
  ungroup() %>%
  mutate(NOr_rev_loc = ifelse(is.na(NOr_revised),
                              NA,
                              paste0(Dorm, Plant, "_", NOr_revised))) %>%
  select(Date, location, NOr_revised, NOr_rev_loc) %>%
  filter(!is.na(NOr_rev_loc))

All_data_v6_N_NOr_counts_unambig_clutches <- All_data_v6_clutch_status_stats_v2 %>%
  right_join(NOr_status_revised_all_unambig_clutches, by = c("Date", "location")) %>%
  group_by(Date, Dorm, Plant, NOr_revised, NOr_rev_loc) %>%
    summarize(N1_NOr = sum(N1, na.rm= TRUE),
              N2_NOr = sum(N2, na.rm= TRUE),
              N5_NOr = sum(N5, na.rm= TRUE),
              N_NOr = sum(N1, N2, N5, na.rm = TRUE),
              A_NOr = sum(n_A, na.rm= TRUE),
              N5plusA_NOr = sum(N5, n_A, na.rm= TRUE)) %>%
  ungroup() # n = 186

All_data_v6_N_NOr_counts_unambig_clutches_summary <- All_data_v6_N_NOr_counts_unambig_clutches %>%
  group_by(Date) %>%
    mutate(mean_N_NOr = mean(N_NOr, na.rm = TRUE),
           mean_N1_NOr = mean(N1_NOr, na.rm = TRUE),
           mean_N2_NOr = mean(N2_NOr, na.rm = TRUE),
           mean_N5_NOr = mean(N5_NOr, na.rm = TRUE),
           mean_A_NOr = mean(A_NOr, na.rm = TRUE),
           mean_N5plusA_NOr = mean(N5plusA_NOr, na.rm = TRUE),
           All_N_NOr = sum(N_NOr),
           All_A_NOr = sum(A_NOr)) %>%
  ungroup() %>%
  group_by(Date, Dorm) %>%
    mutate(All_N_NOr_Dorm = sum(N_NOr),
           All_N1_NOr_Dorm = sum(N1_NOr),
           All_N2_NOr_Dorm = sum(N2_NOr),
           All_N5_NOr_Dorm = sum(N5_NOr),
           All_A_NOr_Dorm = sum(sum(A_NOr))) %>%
  ungroup()

All_data_v6_all_nymphs_unambig_clutches <- All_data_v6_N_NOr_counts_unambig_clutches %>%
  group_by(Date) %>%
    summarize(All_N_NOr = sum(N_NOr)) %>%
  ungroup()

All_data_v6_max_N_NOr_counts_unambig_clutches <-
  All_data_v6_N_NOr_counts_unambig_clutches %>%
  group_by(Dorm, Plant, NOr_revised, NOr_rev_loc) %>%
    summarize(max_N_NOr = max(N_NOr)) %>%
  ungroup()

clutch_status_stats_v1_unambig_clutches_hatch_success <- clutch_status_stats_v1 %>%
  separate(location, c("DormPlant", "Leaf"), sep = "_", remove = FALSE) %>%
  mutate(Dorm = substr(DormPlant, 1, 1),
         Plant = substring(DormPlant, 2)) %>%
  left_join(distinct(clutch_status_guarding_summary_v2, location, never_hatched)) %>%
  right_join(All_data_v6_max_N_NOr_counts_unambig_clutches, by = c("Dorm", "Plant", "Leaf" = "NOr_revised")) %>%
  # why did I originally say to make hatch_success = NA if never_hatched == TRUE?
  mutate(hatch_success = ifelse(!is.na(initiation_date) & is.na(hatch_date),
                                0,
                                max_N_NOr/max_egg_num))
```

## 7.4 - Flag and resolve ambiguous nymph origins

```{r}
NOr_status_resolveable_all_unambig_clutches <- All_data_v6_clutch_status_stats_v2 %>% 
  select(Date, Dorm, Plant, Leaf, location, 
         egg_num, N1, N2, N5, n_A, nymph_origin, initiation_date, hatch_date,
         any_NOr_ambig_loc, Omit_for_abandon_time, never_hatched,
         n_clutches_on_plant) %>%
  # keep only entries with some nymphs or adults on a leaf
  # OR entries for leaves where clutches have been initiated
  group_by(row_number()) %>%
  filter(!is.na(Leaf) & any(N1 > 0, N2 > 0, N5 > 0, n_A > 0) | 
           Date >= initiation_date) %>% # nrow = 17944
  # flag entries where the nymph origin is text (typically ambiguous, e.g. "8 or 9")
  mutate(NOr_ambig_entry = !is.na(nymph_origin) & is.na(as.numeric(nymph_origin))) %>% 
  ungroup() %>%
  # get rid of duplicate rows from multiple females in the same location
  distinct(Date, location, .keep_all = TRUE) %>% # nrow = 15674
  group_by(Dorm, Plant) %>%
    # filter out any plants with a clutch that was omitted for abandon time
    filter(!any(!is.na(Omit_for_abandon_time))) %>%
    # flag plants with any ambiguous entries (e.g. "8 or 9") skip for checking
    mutate(NOr_ambig_plant = any(NOr_ambig_entry)) %>%
  ungroup()

NOr_status_resolveable_mult_clutches <- NOr_status_resolveable_all_unambig_clutches %>%
  group_by(Dorm, Plant) %>%
    # plants with mutliple clutches only
    filter(any(n_clutches_on_plant > 1)) %>%
  ungroup() %>%
  group_by(row_number()) %>%
    # flag for manual checking entries with any nymphs/adults, missing nymph origin, 
    # but not flagged for ambiguous origin in the field.
    # these are the most likely to be easily resolved with minimal work
    mutate(revise_NOr = ifelse(any(N1 > 0, N2 > 0, N5 > 0, n_A > 0) &
             !NOr_ambig_plant,
           "MANUAL CHECK",
           NA)) %>% 
  ungroup()
```

## 7.5 - Fixing the omission of clutches that never had first instar nymphs

```{r}
NOr_status_revised_all_unambig_clutches_v2 <- NOr_status_resolveable_all_unambig_clutches %>%
  left_join(NOr_manual_revisions, by = c("Date", "location")) %>%
  group_by(Dorm, Plant) %>%
    mutate(lone_hatcher_leaf = ifelse(length(unique(Leaf[!is.na(initiation_date)])) == 1,
                                 unique(Leaf[!is.na(initiation_date)]),
                                 NA),
           NOr_revised = ifelse(!is.na(NOr_revised_manual), # if I manually annotated a revised location
                                       NOr_revised_manual, # make manual location the new nymph origin
                                ifelse(any(!is.na(lone_hatcher_leaf)), # if only one clutch hatched on the plant
                                       lone_hatcher_leaf, # make that clutch the new nymph origin
                                       # if no clutches hatched on the plant, NOr_revised = Leaf
                                       # for leaves with eggs
                                       ifelse(!any(!is.na(hatch_date)) & !is.na(egg_num), # if no clutches hatched on the plant (but there are eggs)
                                              Leaf, # make the given leaf the new nymph origin
                                              NA)))) %>% # BUT THIS STILL EXCLUDES CLUTCHES THAT DIDN'T HATCH ON A PLANT WITH SOME CLUTCHES THAT DID HATCH
  ungroup() %>%  
  group_by(location) %>%
    mutate(NOr_revised = ifelse(any(NOr_revised == Leaf) & is.na(NOr_revised), # DOES THIS INCLUDE UNHATCHED CLUTCHES? NO
                                Leaf,#unique(NOr_revised[!is.na(NOr_revised)]),
                                NOr_revised),
           NOr_revised = ifelse(any(is.na(NOr_revised)) & # if no revised nymph origin yet
                                                  any(!is.na(egg_num)&egg_num>0 & !any(!is.na(N1)&N1>0)), # and N1 never on clutch (not hatched or spuriously visited by N1)
                                Leaf, # make this leaf the new nymph origin (so we include unhatched clutches)
                                NOr_revised)) %>% 
  ungroup() %>%
  mutate(NOr_rev_loc = ifelse(is.na(NOr_revised),
                              NA,
                              paste0(Dorm, Plant, "_", NOr_revised))) %>%
  select(Date, location, NOr_revised, NOr_rev_loc) %>%
  filter(!is.na(NOr_rev_loc))

All_data_v6_N_NOr_counts_unambig_clutches_v2 <- All_data_v6_clutch_status_stats_v2 %>%
  right_join(NOr_status_revised_all_unambig_clutches_v2, by = c("Date", "location")) %>%
  group_by(Date, Dorm, Plant, NOr_revised, NOr_rev_loc) %>%
    summarize(N1_NOr = sum(N1, na.rm= TRUE),
              N2_NOr = sum(N2, na.rm= TRUE),
              N5_NOr = sum(N5, na.rm= TRUE),
              N_NOr = sum(N1, N2, N5, na.rm = TRUE),
              A_NOr = sum(n_A, na.rm= TRUE),
              N5plusA_NOr = sum(N5, n_A, na.rm= TRUE)) %>%
  ungroup()

All_data_v6_N_NOr_counts_unambig_clutches_summary_v2 <- All_data_v6_N_NOr_counts_unambig_clutches_v2 %>%
  group_by(Date) %>%
    mutate(mean_N_NOr = mean(N_NOr, na.rm = TRUE),
           mean_N1_NOr = mean(N1_NOr, na.rm = TRUE),
           mean_N2_NOr = mean(N2_NOr, na.rm = TRUE),
           mean_N5_NOr = mean(N5_NOr, na.rm = TRUE),
           mean_A_NOr = mean(A_NOr, na.rm = TRUE),
           mean_N5plusA_NOr = mean(N5plusA_NOr, na.rm = TRUE),
           All_N_NOr = sum(N_NOr),
           All_A_NOr = sum(A_NOr)) %>%
  ungroup() %>%
  group_by(Date, Dorm) %>%
    mutate(All_N_NOr_Dorm = sum(N_NOr),
           All_N1_NOr_Dorm = sum(N1_NOr),
           All_N2_NOr_Dorm = sum(N2_NOr),
           All_N5_NOr_Dorm = sum(N5_NOr),
           All_A_NOr_Dorm = sum(sum(A_NOr))) %>%
  ungroup()
```

# 8 - Calculate hatch success and nymph survivorship

Calculate hatch success (strict sense; max N1 / max eggs) and nymph survivorship (max N5+A / max eggs AND/OR max N5+A / max N1)

- Filtered out clutches with outlier numbers of N1 or N5 etc at early or late clutch ages, which were assumed to be cases of immigration that weren't caught during manual nymph origin filtering
```{r}
# Filter to only include curated list of clutches with unambiguous nymph counts after accounting for immigration to non-natal leaves and then sum all nymphs (including emigrants) for each clutch
# I also added clutch_age variables to help identify clutches that were started too close to the end of sampling to be useful
clutch_unambig_NOr_revised_ENAa_summary_v2 <- All_data_v6_clutch_status_stats_v2 %>%
  right_join(NOr_status_revised_all_unambig_clutches_v2, by = c("Date", "location")) %>%
  select(Date, NOr_rev_loc, location, egg_num, N1, N2, N5, n_A, ants_L) %>%
    # remove any duplicates - I need to figure out why there are dups later but for 4th yr talk
  distinct(Date, NOr_rev_loc, location, egg_num, N1, N2, N5, n_A, ants_L, .keep_all = TRUE) %>%
  mutate(egg_num_NOr = ifelse(NOr_rev_loc == location, egg_num, NA)) %>%
  group_by(Date, NOr_rev_loc) %>%
    summarize(eggs_NOr = first(egg_num_NOr[!is.na(egg_num_NOr)]), # only use eggs from origin leaf!
              N1_NOr = sum(N1, na.rm= TRUE),
              N2_NOr = sum(N2, na.rm= TRUE),
              N5_NOr = sum(N5, na.rm= TRUE),
              N_NOr = sum(N1, N2, N5, na.rm = TRUE),
              A_NOr = sum(n_A, na.rm= TRUE),
              N5toA_NOr = sum(N5, n_A, na.rm= TRUE),
              ants_NOr = sum(ants_L, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(filter(clutch_status_guarding_summary_v2, mom_at_init), by = c("NOr_rev_loc" = "location")) %>%
  mutate(clutch_age1 = Date - initiation_date,
         clutch_age2 = ceiling((Date - initiation_date)/2)*2,
         clutch_age4 = ceiling((Date - initiation_date)/4)*4,
         after_hatch1 = Date - hatch_date,
         after_hatch2 = ceiling((Date - hatch_date)/2)*2,
         after_hatch4 = ceiling((Date - hatch_date)/4)*4) %>%
  select(Date, NOr_rev_loc, clutch_age1, after_hatch1, eggs_NOr, 
         N1_NOr, N2_NOr, N5_NOr, N_NOr, A_NOr, N5toA_NOr, ants_NOr,
         init_fem = Female_ID,
         initiation_date, hatch_date, guard_duration, relative_desert_date,
         clutch_initiated_order, females_removed, ants_removed, 
         clutch_age2, clutch_age4, after_hatch2, after_hatch4) #%>%
  # filter(initiation_date <= as_date("2019-06-10"),
  #        clutch_age4 <= 68)

max_N1_date_v2 <- clutch_unambig_NOr_revised_ENAa_summary_v2 %>%
  group_by(NOr_rev_loc) %>%
  arrange(NOr_rev_loc, desc(N1_NOr), Date) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(NOr_rev_loc, max_N1_date = Date, max_N1_age = clutch_age1)

max_N5_date_v2 <- clutch_unambig_NOr_revised_ENAa_summary_v2 %>%
  group_by(NOr_rev_loc) %>%
  arrange(NOr_rev_loc, desc(N5_NOr), Date) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(NOr_rev_loc, max_N5_date = Date, max_N5_age = clutch_age1)

max_A_date_v2 <- clutch_unambig_NOr_revised_ENAa_summary_v2 %>%
  group_by(NOr_rev_loc) %>%
  arrange(NOr_rev_loc, desc(A_NOr), Date) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(NOr_rev_loc, max_A_date = Date, max_A_age = clutch_age1)

max_N5toA_date_v2 <- clutch_unambig_NOr_revised_ENAa_summary_v2 %>%
  group_by(NOr_rev_loc) %>%
  arrange(NOr_rev_loc, desc(N5toA_NOr), Date) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(NOr_rev_loc, max_N5toA_date = Date, max_N5toA_age = clutch_age1)

max_N5toA_Date_v2 <- clutch_unambig_NOr_revised_ENAa_summary_v2 %>%
  group_by(NOr_rev_loc) %>%
  summarize(max_N1 = max(N1_NOr),
            max_N5toA = max(N5toA_NOr),
            max_clutch_age = max(clutch_age1),
            max_after_hatch = max(after_hatch1),
            max_Date = max(Date),
            hatched = any(!is.na(hatch_date)),
            initiation_date = first(initiation_date)) %>%
  ungroup() %>%
  arrange(desc(max_N5toA)) %>%
  full_join(max_N1_date_v2) %>%
  full_join(max_N5_date_v2) %>%
  full_join(max_A_date_v2) %>%
  full_join(max_N5toA_date_v2)

clutch_unambig_NOr_revised_ENAa_summary_imfilt_v2 <- clutch_unambig_NOr_revised_ENAa_summary_v2 %>%
  mutate(N_imm_likely_d15 = clutch_age1 < 15 & N_NOr > 0,
         N5_imm_likely_d40 = clutch_age1 < 40 & N5_NOr > 0,
         A_imm_likely_d40 = clutch_age1 < 40 & A_NOr > 0) %>%
  group_by(NOr_rev_loc) %>%
    filter(!any(N_imm_likely_d15, N5_imm_likely_d40, A_imm_likely_d40, na.rm = TRUE)) %>% # n = 168
  ungroup()

clutch_unambig_NOr_revised_ENAa_summary_imfilt_success_v2 <-
  clutch_unambig_NOr_revised_ENAa_summary_imfilt_v2 %>%
  group_by(NOr_rev_loc) %>%
    mutate(hatch_success_N1 = max(N1_NOr, na.rm = TRUE)/max(eggs_NOr, na.rm = TRUE),
           survival_rate = max(N5toA_NOr, na.rm = TRUE)/max(eggs_NOr, na.rm = TRUE),
           prop_E2Ni = max(N1_NOr, na.rm = TRUE)/max(eggs_NOr, na.rm = TRUE),
           prop_E2N = max(N_NOr, na.rm = TRUE)/max(eggs_NOr, na.rm = TRUE),
           prop_Ni2Nv = max(N5toA_NOr, na.rm = TRUE)/max(N1_NOr, na.rm = TRUE),
           prop_E2Nv = max(N5toA_NOr, na.rm = TRUE)/max(eggs_NOr, na.rm = TRUE),
           n_eggs = max(eggs_NOr, na.rm = TRUE),
           n_hatched = max(N1_NOr, na.rm = TRUE),
           n_N = max(N_NOr, na.rm = TRUE),
           n_survived = max(N5toA_NOr, na.rm = TRUE)) %>%
  ungroup()

clutch_unambig_NOr_revised_ENAa_summary_imfilt_success_summary_v2 <- clutch_unambig_NOr_revised_ENAa_summary_imfilt_success_v2 %>%
  distinct(NOr_rev_loc, n_eggs, n_hatched, n_survived, n_N,
           prop_E2Ni, prop_Ni2Nv, prop_E2Nv, prop_E2N, hatch_success_N1, survival_rate,
           initiation_date, hatch_date, females_removed, ants_removed) 
```

# 9 - Ant counts on clutch

```{r}
theme_publilia <- function(){
  theme_bw() +
    theme(text = element_text(family = "Arial"),
          axis.text = element_text(size = 12), 
          axis.title = element_text(size = 14),
          axis.line.x = element_line(color="black"), 
          axis.line.y = element_line(color="black"),
          panel.border = element_blank(),
          panel.grid.major.x = element_blank(),                                          
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),  
          plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = , "cm"),
          plot.title = element_text(size = 18, vjust = 1, hjust = 0.5, face = "bold"),
          legend.text = element_text(size = 10),          
          legend.title = element_blank(),                              
          # legend.position = c(0.90, 0.90), 
          legend.key = element_blank(),
          legend.background = element_rect(color = "black", 
                                           fill = "transparent", 
                                           size = 2, linetype = "blank"))
}
```

## 9.0 - Get visited_1st_4d_data

```{r}
strats_data <- left_join(clutch_status_guarding_summary_v2, last_day_fID)

visited_1st_4d_data <- strats_data %>%
  group_by(location) %>%
    mutate(visited_1st_4d = any(!mom_at_init & as.numeric(visit_starts-initiation_date)<=4)) %>%
  ungroup() %>%
  mutate(`visited in 4 days` = ifelse(visited_1st_4d | num_at_init>1,
                                      "Visited", "Not visited"),
         ants = ifelse(ants_removed, "Ants removed", "Control"),
         est_hatch = as_date(ifelse(is.na(hatch_date),
                                 initiation_date + 20,
                                 hatch_date)),
         est_rel2h = visit_ends-est_hatch)
```

```{r}
visited_1st_4d_data_tb <- visited_1st_4d_data %>%
  # 1) Longest initiator
  mutate(initL = init_first_dur_rank == 1) %>% 
  # 2) Shortest initiator
  #  i) break ties by picking marked females over unmarked
  group_by(location, init_first_dur_shortest) %>% 
    mutate(initS_tmp = ifelse(n()>1, !grepl("Uf", Female_ID), TRUE)) %>% 
  ungroup() %>%
  #  ii) finally picking at random (only one case, identical predictor vars)
  group_by(location, init_first_dur_shortest, initS_tmp) %>%
    mutate(initS = init_first_dur_shortest & initS_tmp & row_number()==1) %>%
  ungroup()
```

## 9.1 - A small fraction of ants_removed clutches were visited by ants, and vice versa
```{r}
ant_counts <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  group_by(location) %>%
    mutate(b4h=Date>=initiation_date&Date<est_hatch,
           afh=Date>=est_hatch) %>%
    summarize(ants_removed=first(ants_removed),
              females_removed=first(females_removed),
              n_antsP=sum(ants_P,na.rm=T),
              ever_antsP=n_antsP>0,
              mean_antsP=mean(ants_P,na.rm=T),
              med_antsP=median(ants_P,na.rm=T),

              n_antsP_b4h=sum(ants_P&b4h),
              ever_antsP_b4h=n_antsP_b4h>0,
              mean_antsP_b4h=mean(ifelse(b4h,ants_P,NA),na.rm=T),
              med_antsP_b4h=median(ifelse(b4h,ants_P,NA),na.rm=T),

              n_antsP_afh=sum(ants_P&afh),
              ever_antsP_afh=n_antsP_afh>0,
              mean_antsP_afh=mean(ifelse(afh,ants_P,NA),na.rm=T),
              med_antsP_afh=median(ifelse(afh,ants_P,NA),na.rm=T)) %>%
  ungroup() %>%
  mutate(Dorm=substr(location,1,1),
         ant_treatment=ifelse(ants_removed,"Ants removed","Control"))

ggplot(data=ant_counts,aes(x=Dorm,y=n_antsP, fill=ant_treatment)) +
  geom_boxplot()

ggplot(data=ant_counts,aes(x=Dorm,y=mean_antsP, fill=ant_treatment)) +
  geom_boxplot()

ggplot(data=ant_counts,aes(x=Dorm,y=mean_antsP_b4h, fill=ant_treatment)) +
  geom_boxplot()

ggplot(data=ant_counts,aes(x=Dorm,y=mean_antsP_afh, fill=ant_treatment)) +
  geom_boxplot()

# 9.2 - Threshold for number of days ants were on the plant
Should I be more lenient with my threshold, so I can have more even sample sizes between "ants" and "no ants"?
- ants_ever = n_days_ants>0: F = 83, T = 143
- antsP_gt2d = n_days_ants>2: F = 123, T = 103 * most even sample size
- antsP_gt3d = n_days_ants>3: F = 133, T = 90
```{r}
ant_n_days <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  mutate(ant_treatment=ifelse(ants_removed,"Ants removed","Control"),
         Dorm=substr(location,1,1)) %>%
  group_by(Date, location) %>%
    summarize(any_antsP_day = any(ants_P>0, na.rm=T),
              females_removed = first(females_removed),
              ant_treatment = first(ant_treatment),
              Dorm = first(Dorm)) %>%
  ungroup() %>%
  group_by(location) %>%
    summarize(n_days_ants = sum(as.numeric(any_antsP_day)),
              females_removed = first(females_removed),
              ant_treatment = first(ant_treatment),
              Dorm = first(Dorm))%>%
  ungroup() %>%
  mutate(antsP_gt2d = !is.na(n_days_ants)&n_days_ants>2,
         antsP_gt3d = !is.na(n_days_ants)&n_days_ants>3,
         ants_everP = is.na(n_days_ants)|n_days_ants>0)

ant_n_days %>%
  count(females_removed, ants_everP)

ant_n_days %>%
  count(females_removed, antsP_gt2d)

ant_n_days %>%
  count(females_removed, antsP_gt3d)
```

# 10 - Estimate clutch size more robust to outliers: median top 5 egg counts

## 10.1 - Generate data for fecundity modeling

```{r}
fec_model_data <- visited_1st_4d_data_tb %>%
  filter(!is.na(guard_duration)) %>%
  group_by(location) %>%
  mutate(eggs_ever_visited = any(!mom_at_init & visit_ends <= est_hatch)) %>%
  ungroup() %>%
  select(Female_ID, location, guard_duration, est_rel2h, initiation_date, est_hatch,
         mom_at_init, clutch_initiated_order, initL, initS,
         visit_starts, visit_ends, last_day_fID,
         num_at_init, n_clutches_on_plant, `visited in 4 days`, eggs_ever_visited, ants,
         females_removed, ants_removed, 
         max_egg_num, egg_fate, egg_fate_date)
```

## 10.2 - Calculate median of top 5 egg counts in lieu of max egg counts
Taking the median (or mean) of the top 5 egg_num counts yields an estimate of max egg num that is more robust to outliers/miscounts
```{r}
tmp_t5eggs_data <- clutch_status_init_hatch_labels_v1 %>%
  mutate(est_hatch = as_date(ifelse(is.na(hatch_date),
                                 initiation_date + 20,
                                 hatch_date)),
         clutch_age = as.numeric(Date-initiation_date)) %>%
  # remove duplicate egg counts due to multiple females at a location on a day
  distinct(location, Date, initiation_date, est_hatch, clutch_age,
           egg_num, egg_len) %>%
  group_by(location) %>%
    arrange(desc(egg_num)) %>%
    mutate(top5egg_counts = rank(desc(egg_num), ties.method = "first") <= 5,
           top5egg_lens = rank(desc(egg_len), ties.method = "first") <= 5) %>%
    #filter(location == "A13_11") %>%
    select(location, Date, initiation_date, est_hatch, clutch_age,
           egg_num, egg_len, top5egg_counts, top5egg_lens) %>%
    mutate(med_t5eggs = median(ifelse(top5egg_counts,egg_num,NA), na.rm=T),
           mean_t5eggs = mean(ifelse(top5egg_counts,egg_num,NA), na.rm=T),
           max_n_eggs = max(ifelse(top5egg_counts,egg_num,NA), na.rm=T),
           med_l_t5eggs = median(ifelse(top5egg_lens,egg_len,NA), na.rm=T),
           mean_l_t5eggs = mean(ifelse(top5egg_lens,egg_len,NA), na.rm=T),
           max_l_eggs = max(ifelse(top5egg_lens,egg_len,NA), na.rm=T)) %>%
  ungroup()

top5eggs_summary <- tmp_t5eggs_data %>%
  group_by(location) %>%
    summarize(med_t5eggs = first(med_t5eggs),
              mean_t5eggs = first(mean_t5eggs),
              max_n_eggs = first(max_n_eggs),
              med_l_t5eggs = first(med_l_t5eggs),
              mean_l_t5eggs = first(mean_l_t5eggs),
              max_l_eggs = first(max_l_eggs)) %>%
  ungroup()
```

# 11 - Calculate alternative measures of maternal tenure

`The sum of all females’ egg-guarding durations was more predictive of final clutch size than egg guarding duration of the initiator alone (Table S1 with alternative models and their AICs), showing that females lay eggs while visiting existing clutches.`
